\section{\ac{ism} Band Regulation}\label{sec:ISMBandRegulation}
The industrial, scientific and medical (\ac{ism}) bands are portions of the radio spectrum, which can be used without a license, subject to local regulatory standards. These standards vary around the world but often define maximum power outputs, duty cycles, and bandwidths. Though limits can be problematic, they help reduce the chance of internal and external system interference. Some of the most stringent regulations are in Europe, where they are controlled firstly by the European Telecommunications Standards Institute (the \ac{etsi}\footnote{ETSI, EU, https://www.etsi.org}), and then by country specific authorities. The United States' regulations are managed by the Federal Communications Commission (the \ac{fcc}\footnote{FCC, US, https://www.fcc.gov}), with many other countries following their example. 

Sub-1GHz \ac{lora} hardware operates around the 433MHz and 900MHz bands, but as the former is very heavily regulated for non-remote controls in the US \cite{3YP:FCC_433}, only the 900MHz area is considered viable to the use case. It should be noted however that the specific frequencies used will still differ between region, as is demonstrated by \ac{lorawan}'s use of the 863-870 MHz and 902-928 MHz bands for Europe and the US respectively \cite{3YP:LORAWAN_REGIONAL_PARAMS}. A comparison of the regulations can be seen in Table \ref{tab:ISMRegions}. As \ac{etsi} limits vary heavily on a by band basis, they are broken down further in Table \ref{tab:ETSIBands}. An overview of each regulation type follows.
 
 \begin{table}[H]
\centering\small
\caption[900MHz regional regulation comparison]{Regional regulation comparison for 900MHz band radio \cite{3YP:FCC_900, 3YP:ETSI_HARMONISED_REG}.}
\label{tab:ISMRegions}
\renewcommand*{\arraystretch}{1.1}
\begin{tabular}{l|L{45mm}L{45mm}}
    \toprule
    & \textbf{\ac{fcc}} & \textbf{\ac{etsi}}  \\
    \midrule\addlinespace
    \textbf{Band} & 902--928MHz & 863--870MHz \\
    \textbf{\ac{eirp}} & {36dBm \newline \textit{  30dBm transmit power}} & {0.25MHz @ 27dBm \newline 6.40MHz @ $\leq$ 14dBm} \\
    \textbf{Duty Cycle} & None & 0.1\%--10\% \\
    \textbf{Bandwidth} & 26 MHz & 6.65MHz \\
    \textbf{Narrowband} & 400ms airtime per transmit & None\\
    \textbf{\ac{css}} & None & Varies  \\
    \addlinespace\bottomrule
\end{tabular}
\end{table}

Duty cycle limits greatly reduce the amount of airtime a radio is allowed. For example, in Band h1.3, there is a 1\% duty cycle, which indicates a maximum of 36 seconds of airtime, over all the band's channels, over a rolling one hour period. Duty cycles are considered within a band so a multi-band implementation may obey multiple duty cycles separately. Alternatively, the polite spectrum access (\ac{psa}) policy can be used; this is a defined regulation for listen before talk (LBT) implementations \ac{psa} allows airtime of up to 100 seconds per 200kHz of spectrum, per hour, regardless of duty cycle \cite{3YP:ETSI_PSA}. It does however require a clear channel assessment to be carried out before every transmission, this would be independent of \ac{lora}'s \ac{cad}.
 
\begin{table}
\centering\small
\caption[\ac{etsi} 868MHz sub-band breakdown]{\ac{etsi} 868MHz sub-bands for short range devices (adapted from \cite{3YP:ETSI_HARMONISED_REG}). Only bands relevant for \ac{css} are included. Band numbers correspond to CEPT-ERC-REC 70-03 definitions \cite{3YP:CEPT_ERC_REC}}
\label{tab:ETSIBands}
\renewcommand*{\arraystretch}{1.1}
\begin{tabular}{c|ccccc}
    \toprule
    \textbf{Band} & \textbf{Frequency} (MHz) & \textbf{\ac{eirp}} (dBm)  & \textbf{Duty Cycle} & \textbf{Max \ac{bw}} (kHz) \\
    \midrule\addlinespace
    \textbf{h1.2} & 863.00--870.00 & 14 & 0.1\% (or PSA) & 7000 \\
    \textbf{h1.3} & 865.00--868.00 & 14 & 1\% (or PSA) & 300 \\
    \textbf{h1.4} & 868.00--868.60 & 14 & 1\% (or PSA) & 600 \\
    \textbf{h1.5} & 868.70--869.20 & 14 & 0.1\% (or PSA) & 500 \\
    \textbf{h1.6} & 869.40--869.65 & 27 & 10\% (or PSA) & 250 \\
    \textbf{h1.7a} & 869.70--870.00 & 7 & None & 300 \\
    \textbf{h1.7b} & 869.70--870.00 & 14 & 1\% (or PSA) & 300 \\   
    \addlinespace\bottomrule
\end{tabular}
\end{table}

Regulations consider power as equivalent isotropically radiated power (\ac{eirp}). This value is directly related to radio transmission power ($P_t$) and can be calculated as $EIRP = P_t - L + G$ where $L$ is cable loss and $G$ is antenna gain. The latter occurring from transmit power being concentrated into a smaller area, all antenna will have some form of gain as isotropic antennas are only hypothetical. For example, a radio operating in Band h1.3, using a typical omnidirectional antenna that has 3dBi of gain, can only transmit at 11dBm if no cable loss occurs. Realistically, some cable loss will occur but this must be considered on an implementation by implementation basis. Directional antenna compound these issues due to their high-gains and the relatively low \ac{eirp} limits.

The \ac{etsi} regulations are the limiting factor in transmit power, duty cycle and overall available bandwidth. \ac{lora} being a \ac{css} signal means \ac{fcc} narrowband limitations need not be a concern. Under this consideration, the \ac{etsi} regulations will be used as the worst-fit scenario from here on in. When considering national regulation, it is common that either all bands are implemented, or none at all \cite{3YP:CEPT_ERC_REC}. Of the ETSI bands, h1.3, h1.4 and h1.6 are of most interest. h1.3 and h1.4 giving a balanced offering of bandwidth, \ac{eirp} and duty cycle. Whilst h1.6 offers a far greater duty cycle and \ac{eirp} but limited bandwidth; a further regulatory limit means only allows a single wide-band channel can operate within this band. The maximum number of possible \ac{lora} channels in each of these bands can be seen in Table \ref{tab:ETSILoraChannels}.

\begin{table}[H]
\centering\small
\caption[Maximum channel breakdown for \ac{lora}]{Maximum channel count breakdown for relevant \ac{etsi} bands, calculated for \ac{lora}'s most common operating bandwidths. Calculated assuming channel spacing is 120\%  of the bandwidth to avoid inter-channel interference.}
\label{tab:ETSILoraChannels}
\renewcommand*{\arraystretch}{1.1}
\begin{tabular}{c|ccc}
    \toprule
    & \multicolumn{3}{c}{\textbf{Channel \ac{bw}}}	\\
    \textbf{Band} & \textbf{125kHz} & \textbf{250kHz} & \textbf{500kHz}\\
    \midrule\addlinespace
    \textbf{h1.3} & 19 & 8 & N/A \\
    \textbf{h1.4} & 4 & 2 & 1 \\
    \textbf{h1.6} & 1 & 1 & 0 \\
    \addlinespace\bottomrule
\end{tabular}
\end{table}
