%% ----------------------------------------------------------------
%% Project.tex
%% ---------------------------------------------------------------- 
\documentclass{ecsproject}     
\usepackage[backend=bibtex8, style=ieee]{biblatex}
\usepackage{acro}
\usepackage[singlelinecheck=false]{caption}
\usepackage{booktabs}
\usepackage{float}
\usepackage{makecell}
\usepackage{subfig}
\usepackage{pgfgantt}
\usepackage{longtable}
\usepackage[shortlabels]{enumitem}
\usepackage{array}
\usepackage{multirow}
\usepackage{colortbl}
\usepackage[math]{cellspace}
\usepackage{amsmath}
\usepackage[edges]{forest}

\newcommand{\PreserveBackslash}[1]{\let\temp=\\#1\let\\=\temp}
\newcolumntype{C}[1]{>{\PreserveBackslash\centering}m{#1}}
\newcolumntype{R}[1]{>{\PreserveBackslash\raggedleft}m{#1}}
\newcolumntype{L}[1]{>{\PreserveBackslash\raggedright}m{#1}}
\newcommand{\centered}[1]{\begin{tabular}{l} #1 \end{tabular}}

\input{Definitions}
\input{Abbreviations}

\graphicspath{{../Figures/}}
\hypersetup{colorlinks=false}
\acsetup{first-style=short}
\bibliography{citations}

%% ----------------------------------------------------------------
\begin{document}
\frontmatter
\title      {LoRa Communications for Sparse Robot Swarms}
\authors    {\texorpdfstring
             {\href{mailto:dsj1n15@soton.ac.uk}{David Jones}}
             {David Jones}
            }
\addresses  {\groupname\\\deptname\\\univname}
\date       {\today}
\subject    {}
\keywords   {}
\supervisor {Dr Klaus-Peter Zauner}
\examiner   {Prof. Kirk Martinez}
\degree     {MEng Computer Science}
\maketitle

\begin{abstract}

Swarms are mobile multi-robot systems that communicate in a decentralised manner to achieve their goals. One potential application of swarms is search and rescue in remote environments (where cellular infrastructure is unavailable). When robot cost is significant, robots will be deployed sparsely meaning typical point-to-point high-data-rate technologies like Wi-Fi cannot be applied. This paper identifies the long-range, low-data-rate technology of LoRa to serve this scenario. 

The technology's performance was first verified through real-world radio measurements for potential swarm scenarios. This data was built into simulation models. On which, three MAC protocols with the aim of distributing local-interest data to physically local neighbours, were tested. Protocols were constrained by regional (ETSI) regulations and the physical performance of LoRa. Two methods utilised common approaches: ALOHA and CSMA. The third, LLBP, was bespoke, and attempted to exploit hardware parameters and regional regulations. 

Given that LoRa is typically deployed in sensor networks, leading to radio testing literature using highly elevated antenna,  the physical radio data collected is novel. Ranges of 500m were achieved in-forest, with 1600m out-of-forest, this is significantly less than perfect LoRa deployments which frequently achieve 10km. Of the three protocols, ALOHA and CSMA performed equally well with LLBP unable to exploit LoRa as intended. All protocols were heavily stifled by regulation with maximum throughput of 15KB per hour, per radio achieved. Whether this is enough for swarm applications must be assessed on an implementation basis.
 
\end{abstract}
\input{StatementOfOrginality}
\newpage{}
\acknowledgements{
This work would not have been possible without Klaus Peter-Zauner's weekly guidance and continuous support. I am also thankful to Kirk Martinez for his invaluable suggestions on simulation methods, which kept this project progressing. Most of all I would like to thank any family and friends who kept me company through the many hours of data collection, no matter the weather!
}

\tableofcontents
\listoffigures
\listoftables
\listofacronyms
%\input{Symbols}

\mainmatter

%% ----------------------------------------------------------------
\include{Introduction}
\include{Background}
\include{PHYTesting}
\include{Simulator}
\include{Protocol}
\include{Conclusion}
\include{Retrospective}

\printbibliography

\appendix
\include{Appendix}
\backmatter

\end{document}
%% ----------------------------------------------------------------
